---
title: "Biden’s Approval Margins"
author: "Alice Chen"
date: "2022-09-12"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```

```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
```

fivethirtyeight.com has detailed data on [all polls that track the president\'s approval](https://projects.fivethirtyeight.com/biden-approval-ratings)

```{r read_GDP_data}

# Import approval polls data directly off fivethirtyeight website
approval_polllist <- read_csv('https://projects.fivethirtyeight.com/biden-approval-data/approval_polllist.csv') 

glimpse(approval_polllist)

```

```{r reshape_GDP_data}

# Use `lubridate` to fix dates, as they are given as characters.

#Change eddate to date type from character
approval_polllist$enddate <- mdy(approval_polllist$enddate)
```

## 2.1 Create a plot

What we would like to do is to calculate the average net approval rate (approve- disapprove) for each week since he got into office. Later, plot the net approval for each week in 2022, along with its 95% confidence interval. There are various dates given for each poll, please use `enddate`, i.e., the date the poll ended.

```{r gdp1}
#Create line plot with confidence intervals
approval_polllist %>% 
  filter(lubridate::year(enddate)==2022) %>% 
  
  #Calculating net approval rate and week of the year
  mutate(net_approval_rate = approve-disapprove, week_of_year = week(enddate)) %>% 
  
  #calculating summary stats & 95% confidence interval values for each week + each subgroup
  group_by(week_of_year, subgroup) %>% 
  summarise(mean_net_approval_rate = mean(net_approval_rate),
            count = n(),
            sd = sd(net_approval_rate),
            t_critical = qt(0.975, count-1),
            se_net_approval_rate = sd / sqrt(count),
            margin_of_error = t_critical * se_net_approval_rate,
            lower_approval_rate = mean_net_approval_rate - margin_of_error,
            higher_approval_rate = mean_net_approval_rate + margin_of_error) %>% 
  
  #order by week of year
  arrange(week_of_year) %>% 
  
  #plot line graph
  ggplot(aes(x=week_of_year , y=mean_net_approval_rate, color = subgroup)) +
  geom_line() +
  
  #Add confidence interval to line plots
  geom_ribbon(aes(ymin = lower_approval_rate, ymax = higher_approval_rate), 
              alpha = 0.1,
              fill = "black") +
  facet_wrap(~subgroup, nrow = 3, strip.position = "right") +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "Biden's Net Approval Ratings in 2022", 
       subtitle = "Weekly Data, Approve-Disapprove, %",
       x = "Week in 2022",
       y="")
```
